Coherent Hyper-Raman Four Wave Mixing Vibrational Spectroscopies

### 2024-07-09 paper.tex
fixed more in the singly resonant section
need to focus on the doubly res section now (make figures particuarly..)

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics.
Three wave mixing vibrational spectroscopies are similarly used to understand the spectroscopy and dynamics of interfacial species.
Most of these methods rely on infrared or Raman transitions to generate output. 
Implementing a nonlinear spectroscopy involving hyper-Raman transitions, first appearing under third order perturbation theory, can provide a useful analogue to infrared and Raman based nonlinear spectroscopies to understand the spectroscopy and dynamics of isotropic systems.
Hyper Difference Frequency Generation (HDFG) spectroscopy, also known as singly vibrationally enhanced spectroscopy (SIVE), is an underdeveloped, hyper-Raman based four wave mixing vibrational spectroscopy. 
However, the spectroscopic and dynamic properties of HDFG have not been fully explored.
We derive selection rules for singly resonant hyper-DFG and demonstrate it provides a simpler method to extract hyper-Raman polarizabilities ($\beta$).
HDFG output is shown to be similar in magnitude to vibrational sum frequency generation (vSFG). 
We demonstrate the properties and capabilities of HDFG through measuring spectra and coherence dynamics of multiple infrared active modes in CH$_3$CN.
HDFG shows promise as a method to disentangle ultrafast vibrational dynamics in isotropic systems without need for anharmonicities.

@@

\begin{figure}[!htbp]
	\centering
	\includegraphics[width=3.375in]{f1.png}
	\caption{Wave Mixing Energy Level (WMEL) diagrams of difference frequency generation (DFG), spontaneous Raman scattering, hyper difference frequency generation (HDFG) and spontaneous hyper-Raman scattering. \cite{RN286, RN352}
		$\ket{g}$ is the ground state, $\ket{v}$ is an infrared active vibration and $\ket{m}, \ket{n}$ are virtual states.
		Solid and dashed horizontal lines indicate real and virtual states, whereas solid and dotted arrows indicate ket and bra side transitions, respectively. %todo: add Raman to this WMEL! dfg Raman hyperDFG hyperRaman 
		}
	\label{fig:comparisonwmel}
\end{figure}

@@

To employ the Albrecht formalism, we write the states in terms of a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a) \otimes \ket{b}$ for electronic states $\{|a)\}$ and vibrational states $\{\ket{b}\}$. \cite{BornOppenheimer, Albrecht1960}
For simplicity, we choose $\ket{g}, \ket{v}$ to be vibrational states on the ground electronic state $|g)$ and $\ket{m} \rightarrow \ket{m,j}$, $\ket{n} \rightarrow \ket{n,k}$ are virtual states.
For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments ($\vec{\mu}$ is reserved for transitions between states on the ground electronic state). \cite{Ziegler1974}
Following the approach which gave \autoref{sivegamma}, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,j,k} \frac{
		R_{i}^{gv, nk} 
		R_{j}^{nk,mj} 
		R_{k}^{mj,gg} 
		R_{l}^{gv,gg} 
	}{\Delta_{g0,gg}
		\Delta_{nk, mj}
		\Delta_{mj, gg}
	}
\end{equation}
where $R_{i}^{ab,cd}$ is the i$^{th}$ element of $\mel{a,b}{\vec{R}}{c,d}$.\cite{Ziegler1988}
Using our definition of vibronic states, we write $\vec{M}^{ab} = (a|\vec{M}|b)$ so that, for example,
$R_{i}^{gv,ev'} = \mel{v}{M_i^{ge}}{v'}$.\cite{Albrecht1960}
Using these ideas rewrites \autoref{drgamma_notaylor} as
\begin{equation}\label{drgamma_notaylor_1}
	\begin{split}
		\gamma_{ijkl} &= -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
			\mel{v}{M_{i}^{ge}}{v'} 
			\mel{v'}{M_{j}^{em}}{n}
			\mel{n}{M_{k}^{mg} }{v}
			\mel{v}{M_{l}^{gg}}{0}
		}{\Delta_{g0,gv}
			\Delta_{ev', mn}
			\Delta_{mn, g0}	} \\
%	&= 	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
%		\mel{v}{M_{i}^{ge}}{v'} 
%		\mel{v'}{M_{j}^{em}}{n}
%		\mel{n}{M_{k}^{mg} }{v}
%		\mel{v}{\mu_{l}}{0}
%	}{\Delta_{g0,gv}
%		\Delta_{ev', mn}
%		\Delta_{mn, g0}	} \\
	\end{split}
\end{equation}
Chung and Ziegler showed some years ago that expanding $\vec{M}^{ij}$ to $\order{Q}$ as
\begin{equation}
\vec{M}^{ij} = \vec{M}^{ij}_0 + \sum_z \frac{\partial\vec{M}^{ij}}{\partial Q_z}
\end{equation} yields $A, B, C$ coefficients similar to those in the Albrecht formalism of Raman scattering. \cite{Albrecht1961, Ziegler1988}
The $A$ term depends upon Franck-Condon ($\order{Q^0}$) transitions, the $B$ term depends upon Herzberg-Teller ($\order{Q}$) transitions, and the $C$ term depends on Duschinsky ($\order{Q^2}$) transitions. 
Here, we ignore the C term as it depends on one and two photon forbidden transitions which have yet to be observed in spontaneous hyper-Raman scattering. \cite{Ziegler1988}
The full derivation of $A,B$ have been given elsewhere, so only their final forms will be given.
In the discussion that follows, we take $\hbar = e = 1$ for simplicity.

By compiling all the terms that scale with $\order{Q^0}$, i.e. $\vec{M}^{ij}_0$, $A$ is written as
\begin{equation}
	\begin{split}
		A_{ijk} = \sum_{m,n,j,k} M^{gn}_{0,i} 
		M^{nm}_{0,j} 
		M^{mg}_{0,k}
		 \langle v | k \rangle
		 \langle k | j \rangle
		 \langle j | g \rangle \\
	\end{split}
\end{equation}
and $B$ is written as a sum of two terms, $B_1, B_2$, as
\begin{subequations}
	\begin{equation}
		\begin{split}
			B_{1_{ijk}} &= \sum_{m,n,j,k,z} M^{gn}_{0,i} \langle v | k \rangle \lbrack
			 \frac{\partial M^{nm}_{j}}{\partial Q_z} \mel{k}{Q_z}{j} M^{mg}_{0,k} \langle j | g \rangle
			+M^{nm}_{0,j} \langle k | j \rangle \frac{\partial M^{mg}_{k}}{\partial Q_z} \mel{j}{Q_z}{g} \rbrack \\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{ijk}} = \sum_{m,n,j,k,z} \frac{\partial M^{gn}_{i}}{\partial Q_z} \mel{v}{Q_z}{k} M^{nm}_{0,j} 
			M^{mg}_{0,k}
			\langle k | j \rangle
			\langle j | g \rangle
	\end{split}
	\end{equation}
\end{subequations}
where terms such as $\langle a | b \rangle$ are Franck-Condon factors and $\mel{a}{Q_z}{b}$ are Herzberg-Teller coupling factors. %include parameterization of the states?
When $\omega_2$ and $\omega_3$ are nonresonant, the resonance denominators lose their dependence on the vibrational quantum number and the sum can be contracted over ${j, k}$. 
Assuming a normal mode basis, this takes $\sum_{j,k}\langle v | k \rangle \langle k | j \rangle \langle j | g \rangle = \delta_{vg} = 0$, or, A$_{ijk}$ vanishes for singly resonant HDFG. \cite{Neddersen1989}
The B terms do not suffer from this problem.
The Herzberg-Teller matrix elements, where $Q_z$ is a normal mode of the system which is not $\ket{g}$ or $\ket{v}$, allows for coupling between different terms that result in non-zero B terms.
As a result, singly resonant HDFG has significant dependence upon those modes which participate in Herzberg-Teller coupling. 
This likely explains the relative weakness of HDFG in solvents such as hexane. \cite{RN351}
The singly resonant HDFG response can then be approximated as
\begin{equation}
	\gamma_{ijkl} \sim B_{ijk} \mu_l
\end{equation}
or, while a vibrational mode must be infrared active for HDFG output, it must also significantly participate in vibronic coupling, i.e., large B values are a necessity for singly resonant HDFG response.
\subsection{Doubly Resonant}
While HDFG can provide useful information about the properties of molecular vibrations, it is apparent that HDFG can inform on molecular electronic structure by making the hyper-Raman transition resonant with a vibronic state.
DR-HDFG (where a vibrational and vibronic state are resonantly coupled) can provide a tool for measuring vibronic coupling.
This method is analogous to doubly resonant sum frequency generation (DR-SFG). \cite{Shen94}
Many multi-resonant nonlinear spectroscopies have been developed to resolve vibronic coupling in molecular samples. \cite{Carlson1990, Gaynor2017, RN276}
DR-HDFG provides methods to investigate vibronic coupling with only two laser pulses.
Two beam, parametric and non-parametric DR-HDFG processes are diagrammed in \autoref{fig:doubsive}.

@@

Since the hyper-Raman process is stimulated by a two-photon process, a setup using only two lasers can perform a coherent four wave mixing hyper-Raman experiment by investigating $\pm \vec{k}_1 + 2\vec{k}_2$ processes, assuming a two photon absorption from a single pulse stimulates the hyper-Raman process.
The other commonly used two laser CMDS method is sum frequency generation, a $\chi^{(2)}$ technique, whose phasematching is dictated by $\vec{k}_1 + \vec{k}_2$.
The SFG output scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor.
Since SFG is dependent upon transitions that transform like both rank-one and rank-two tensors, this enforces a microscopic constraint that SFG cannot yield output for centrosymmetric species under the electric dipole approximation.
This follows from the presence of an inversion center in centrosymmetric systems, which demands strict parity selection rules. \cite{RN230}
Macroscopically, inversion symmetry in the output polarization eliminates output from centrosymmetric species under the electric dipole approximation.\cite{RN227, RN132}
As a result, the output of SFG is significantly reduced relative to most third order spectroscopies because the output only depends on surface species.
However, since HDFG is dependent upon a hyper-Raman transition, it is useful to compare the relative output of the two processes to identify how HDFG output compares to SFG.

To motivate the application of HDFG spectroscopy, we perform a calculation to compare HDFG and vibrational SFG (vSFG) output.
Absorption effects are neglected for simplicity.
\begin{widetext}
To allow comparison between the two methods, the output polarizations of each process is written as
	\begin{subequations}
		\begin{equation}
			\abs{P^{(2)}_{\text{SFG}}} = N_{surf} F(\omega_1+\omega_2) \abs{\langle \alpha_{ij}\mu_{k} \rangle E_J(\omega_2)E_K(\omega_1)} 
		\end{equation}
		\begin{equation}
			\abs{P^{(3)}_{\text{HDFG}}} = N_{bulk}  F(-\omega_1+\omega_2 +\omega_3) \abs{\langle \beta_{ijk} \mu_{l} \rangle E_J{(\omega_3)}E_K(\omega_2)E_L(-\omega_1)} \ell_{eff}
		\end{equation}
	\end{subequations}
where N$_{bulk}$ is the bulk number density, and N$_{surf}$ is the surface number density (10$^{19}$ m$^{-2}$).\cite{RN133, RN503}	
For liquids with molar masses less than $\sim$100 g mol$^{-1}$ and densities roughly equal to that of H$_2$O$_{(l)}$ at room temperature, N$_{bulk} \sim$ 10$^{28}$ m$^{-3}$.
We take the bulk optical interaction length ($\ell_{eff}$) to be 10 $\mu$m.\cite{RN133} %see chapter 25 of Shen's NLO book... idk something here seems wrong
A normal dispersion curve is assumed so that $F(\omega_1+\omega_2) \approx F(-\omega_1+\omega_2 +\omega_3)$.
To simplify analysis, orientational averaging is ignored (e.g., $\langle \beta_{ijk} \mu_{l} \rangle \approx \beta \mu$) and the input and output fields are taken to be co-polarized, so that
\begin{equation}
		P_{ratio} \equiv \frac{\abs{P^{(3)}_{\text{HDFG}}}}{\abs{P^{(2)}_{\text{SFG}}}} \approx \frac{N_{bulk}}{N_{surf}} \frac{\beta}{\alpha} E(\omega_3) \ell_{eff} \sim 10^4 \frac{\beta}{\alpha} E(\omega_3)\\
\end{equation}
\end{widetext}
Ziegler has noted that for a field of intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_3)$ is largely detuned from electronic resonances. \cite{RN515}
Such an intensity is easily obtained using modern ultrafast sources.
In this limit, $P_\text{ratio} \sim 10$.
Since the intensity ratio scales as $\abs{P_{ratio}}^2$, we see that the HDFG output is somewhat stronger than vibrational SFG, assuming only interfacial contributions in vSFG.
It is important to note that this derivation ignored absorption and orientational averaging effects, which can significantly reduce output from either process.

@@

The steps behind orientational averaging of $\gamma_{ijkl}$, a rank four tensor in the molecular frame, are detailed elsewhere.\cite{Andrews1977, McDonnell2024}
Orientational averaging shows specific polarization schemes isolate linear combinations of different $\beta_{ijk}$ terms. 
Infrared spectra can provide $\vec{\mu}$ values.
Since $\vec{\mu}$ is a rank-one tensor, its three Cartesian components share the simple relationship $\langle {\mu^2_x} \rangle = \langle {\mu^2_y} \rangle = \langle {\mu^2_z} \rangle = \frac{1}{3}\langle {\mu^2} \rangle$. \cite{RN459}
Neglecting rotational dependence (i.e., assuming only vibrational transitions), $\langle {\mu^2} \rangle = \mu^2$, so that $\langle \beta_{ijk} \mu_l \rangle = \langle \beta_{ijk} \rangle \frac{\abs{\mu}}{\sqrt{3}}$.
By introducing the reduced mass coordinate $q_n = \sqrt{\frac{1}{m_n}} Q_n$ and using the expansions of $\beta_{ijk}$ and $\mu_{l}$ to $\order{Q_n}$ found earlier, we see
\begin{equation}\label{betasive}
	\langle \frac{\partial \beta_{ijk}}{\partial q_n} \rangle = \varepsilon_0 \frac{48 \sqrt{3} }{NF}  \frac{\Gamma_{gv} \omega_{vg}}{\rho_{gg}} \frac{\Im(\chi^{(3)}_{IJKL})}{\frac{\partial \mu}{\partial q_n}}
\end{equation}
Since $\abs{\mu}$ values can be readily extracted from FT-IR spectra,\cite{RN412} this shows that HDFG can give quantitative information on the magnitude of $\beta_{ijk}$.
HDFG and vSFG together provide unique methods to extract the microscopic quantities ($\mu, \alpha, \beta$) that drive nonlinear output.

>>>>

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics.
Three wave mixing vibrational spectroscopies are similarly used to understand the spectroscopy and dynamics of interfacial species.
Most of these methods rely on infrared or Raman transitions to generate output. 
Implementing a nonlinear spectroscopy involving hyper-Raman transitions, first appearing under third order perturbation theory, can provide a useful analogue to infrared and Raman based nonlinear spectroscopies to understand the spectroscopy and dynamics of isotropic systems.
Hyper Difference Frequency Generation (HDFG) spectroscopy, also known as singly vibrationally enhanced spectroscopy (SIVE), is an underdeveloped, hyper-Raman based four wave mixing vibrational spectroscopy. 
The spectroscopic and dynamic properties of HDFG have not been fully explored.
We derive selection rules for singly resonant hyper-DFG and demonstrate it provides a simpler method to extract hyper-Raman polarizabilities ($\beta$).
HDFG output is shown to be similar in magnitude to vibrational sum frequency generation (vSFG). 
We demonstrate the properties and capabilities of HDFG through measuring spectra and coherence dynamics of multiple infrared active modes in CH$_3$CN.
HDFG shows promise as a method to disentangle ultrafast vibrational dynamics in isotropic systems without need for anharmonicities.

@@

\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66in]{f1.png}
	\caption{Wave Mixing Energy Level (WMEL) diagrams of difference frequency generation (DFG), spontaneous Raman scattering, hyper difference frequency generation (HDFG) and spontaneous hyper-Raman scattering. \cite{RN286, RN352}
		$\ket{g}$ is the ground state, $\ket{v}$ is an infrared active vibration and $\ket{m}, \ket{n}$ are virtual states.
		Solid and dashed horizontal lines indicate real and virtual states, whereas solid and dotted arrows indicate ket and bra side transitions, respectively. %todo: add Raman to this WMEL! dfg Raman hyperDFG hyperRaman 
		}
	\label{fig:comparisonwmel}
\end{figure*}

@@

To employ the Albrecht formalism, we write the states in terms of a Born-Oppenheimer basis $\ket{a,b}$, where $\ket{a,b} = |a) \otimes \ket{b}$ for electronic states $\{|a)\}$ and vibrational states $\{\ket{b}\}$. \cite{BornOppenheimer, Albrecht1960}
For simplicity, we choose $\ket{g}, \ket{v}$ to be vibrational states on the ground electronic state $|g)$ and $\ket{m} \rightarrow \ket{m,j}$, $\ket{n} \rightarrow \ket{n,k}$ are virtual states.
For consistency with previous reports, we use $\vec{R}$ to denote electric transition dipole moments ($\vec{\mu}$ is reserved for transitions between states on the ground electronic state). \cite{Ziegler1974}
Following the approach which gave \autoref{sivegamma}, we find
\begin{equation}\label{drgamma_notaylor}
	\gamma_{ijkl} = -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,j,k} \frac{
		R_{i}^{gv, nk} 
		R_{j}^{nk,mj} 
		R_{k}^{mj,gg} 
		R_{l}^{gv,gg} 
	}{\Delta_{g0,gg}
		\Delta_{nk, mj}
		\Delta_{mj, gg}
	}
\end{equation}
where $R_{i}^{ab,cd}$ is the i$^{th}$ element of $\mel{a,b}{\vec{R}}{c,d}$.\cite{Ziegler1988}
Using our definition of vibronic states, we write $\vec{M}^{ab} = (a|\vec{M}|b)$ so that, for example,
$R_{i}^{gv,ev'} = \mel{v}{M_i^{ge}}{v'}$.\cite{Albrecht1960}
Using these ideas rewrites \autoref{drgamma_notaylor} as
\begin{equation}\label{drgamma_notaylor_1}
	\begin{split}
		\gamma_{ijkl} &= -\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
			\mel{v}{M_{i}^{ge}}{v'} 
			\mel{v'}{M_{j}^{em}}{n}
			\mel{n}{M_{k}^{mg} }{v}
			\mel{v}{M_{l}^{gg}}{0}
		}{\Delta_{g0,gv}
			\Delta_{ev', mn}
			\Delta_{mn, g0}	} \\
%	&= 	-\frac{1}{\varepsilon_0} \frac{1}{24 \hbar^3} \sum_{m,n,v'} \frac{
%		\mel{v}{M_{i}^{ge}}{v'} 
%		\mel{v'}{M_{j}^{em}}{n}
%		\mel{n}{M_{k}^{mg} }{v}
%		\mel{v}{\mu_{l}}{0}
%	}{\Delta_{g0,gv}
%		\Delta_{ev', mn}
%		\Delta_{mn, g0}	} \\
	\end{split}
\end{equation}
Chung and Ziegler showed some years ago that expanding $\vec{M}^{ij}$ to $\order{Q}$ as
\begin{equation}
\vec{M}^{ij} = \vec{M}^{ij}_0 + \sum_z \frac{\partial\vec{M}^{ij}}{\partial Q_z}
\end{equation} yields $A, B, C$ coefficients similar to those in the Albrecht formalism of Raman scattering, \cite{Albrecht1961, Ziegler1988} so that
\begin{equation}
		\gamma_{ijkl} \sim \frac{(A_{ijk} + B_{ijk} + C_{ijk})\mel{v}{\mu_{l}}{0}} {\Delta_{g0,gv}}
\end{equation}
The $A$ term depends upon Franck-Condon ($\order{Q^0}$) transitions, the $B$ term depends upon Herzberg-Teller ($\order{Q}$) transitions, and the $C$ term depends on $\order{Q^2}$ transitions. 
The C term is suppressed in the following discussion as it depends on one and two photon forbidden transitions which have yet to be observed in spontaneous hyper-Raman scattering. \cite{Ziegler1988}
In the discussion that follows, we take $\hbar = e = 1$.

By compiling all the terms that scale with $\order{Q^0}$, $A$ is written as
\begin{equation}
	\begin{split}
		A_{qrs} = \sum_{m,n,j,k} M^{gn}_{0,q} 
		M^{nm}_{0,r} 
		M^{mg}_{0,s}
		 \langle v | k \rangle
		 \langle k | j \rangle
		 \langle j | g \rangle 
		 \frac{1}{\Delta_{nk,mj} \Delta_{mj, gg}}
		 \\
	\end{split}
\end{equation}
and by compiling all terms that scale with $\order{Q}$, $B$ is written as $B = B_1 + B_2$,\cite{Bonang1992} where
\begin{subequations}
	\begin{equation}
		\begin{split}
			B_{1_{qrs}} &= \sum_{m,n,j,k,z} M^{gn}_{0,q} \langle v | k \rangle \lbrack
			 \frac{\partial M^{nm}_{r}}{\partial Q_z} \mel{k}{Q_z}{j} M^{mg}_{0,s} \langle j | g \rangle
			+M^{nm}_{0,r} \langle k | j \rangle \frac{\partial M^{mg}_{s}}{\partial Q_z} \mel{j}{Q_z}{g} \rbrack \frac{1}{\Delta_{nk,mj} \Delta_{mj, gg}}\\
		\end{split}
	\end{equation}
	\begin{equation}
	\begin{split}
			B_{2_{qrs}} = \sum_{m,n,j,k,z} \frac{\partial M^{gn}_{q}}{\partial Q_z} \mel{v}{Q_z}{k} M^{nm}_{0,r} 
			M^{mg}_{0,s}
			\langle k | j \rangle
			\langle j | g \rangle 
			\frac{1}{\Delta_{nk,mj} \Delta_{mj, gg}}
	\end{split}
	\end{equation}
\end{subequations}
where terms such as $\langle a | b \rangle$ are Franck-Condon factors and $\mel{a}{Q_z}{b}$ are Herzberg-Teller coupling factors. %include parameterization of the states?

When $\omega_2$ and $\omega_3$ are largely detuned from a local resonance, the resonance denominators lose their dependence on vibrational quantum number.
In the case of the $A$ term, the sum can now be contracted over ${j, k}$. 
Assuming a normal mode basis, $A_{ijk} \sim \sum_{j,k}\langle v | k \rangle \langle k | j \rangle \langle j | g \rangle = \delta_{vg} = 0$, or, A$_{ijk}$ vanishes for singly resonant HDFG. \cite{Neddersen1989}
The B terms do not suffer from this problem.
The Herzberg-Teller matrix elements, where $\ket{z} \neq \{\ket{g}, \ket{v}\}$ is a normal mode, allows for non-zero B terms.
As a result, singly resonant HDFG has significant dependence upon those modes which participate in Herzberg-Teller coupling. 
This likely explains the relative weakness of HDFG response from hexane $\nu$(CH) modes. \cite{RN351}
The singly resonant HDFG susceptibility can then be approximated as
\begin{equation}
	\chi^{(3)}_{IJKL} \sim \langle B_{ijk} \mu_l \rangle
\end{equation}
or, while a vibrational mode must be infrared active for HDFG output, large $B$ values are a necessity for singly resonant HDFG response.
\subsection{Doubly Resonant}
While HDFG can provide useful information about the properties of molecular vibrations, it is apparent that HDFG can inform on molecular electronic structure by making the hyper-Raman transition resonant with a vibronic state, i.e. doubly resonant HDFG (DR-HDFG).
DR-HDFG (where a vibrational and vibronic state are resonantly coupled) can provide a tool for measuring vibronic coupling, analogous to fully resonant DFG. \cite{Dick83_1, Shen94}
Many multi-resonant nonlinear spectroscopies have been developed to resolve vibronic coupling in molecular samples. \cite{Carlson1990, Gaynor2017, RN276}
DR-HDFG provides methods to investigate vibronic coupling with only two laser pulses.
Two beam, parametric and non-parametric DR-HDFG processes are diagrammed in \autoref{fig:doubsive}.

@@

Since the hyper-Raman process is stimulated by a two-photon process, a setup using only one optical parametric amplifier can perform a coherent four wave mixing hyper-Raman experiment by investigating $\pm \vec{k}_1 + 2\vec{k}_2$ processes, assuming a two photon absorption from a single pulse stimulates the hyper-Raman process.
The other commonly used two laser CMDS method is sum frequency generation, a $\chi^{(2)}$ technique, whose phasematching is dictated by $\vec{k}_1 + \vec{k}_2$.
The SFG output scales as $\chi^{(2)}_{IJK} \sim \langle \alpha_{ij} \mu_k \rangle$, where $\alpha_{ij}$ is the Raman polarizability tensor.
Since SFG is dependent upon transitions that transform like both rank-one and rank-two tensors, this enforces a microscopic constraint that SFG cannot yield output for centrosymmetric species under the electric dipole approximation.
This follows from the presence of an inversion center in centrosymmetric systems, which demands strict parity selection rules. \cite{RN230}
Macroscopically, inversion symmetry in the output polarization eliminates output from centrosymmetric species under the electric dipole approximation.\cite{RN227, RN132}
As a result, the output of SFG is significantly reduced relative to most third order spectroscopies because the output only depends on surface species.
However, since HDFG is dependent upon a hyper-Raman transition, it is useful to compare the relative output of the two processes to identify how HDFG output compares to SFG.

To motivate the application of HDFG spectroscopy, we perform a calculation to compare HDFG and vibrational SFG (vSFG) output.
Absorption effects are neglected for simplicity.
The output polarizations of each process is written as
	\begin{subequations}
		\begin{equation}
			\abs{P^{(2)}_{\text{SFG}}} = N_{surf} F(\omega_1+\omega_2) \abs{\langle \alpha_{ij}\mu_{k} \rangle E_J(\omega_2)E_K(\omega_1)} 
		\end{equation}
		\begin{equation}
			\abs{P^{(3)}_{\text{HDFG}}} = N_{bulk}  F(-\omega_1+\omega_2 +\omega_3) \abs{\langle \beta_{ijk} \mu_{l} \rangle E_J{(\omega_3)}E_K(\omega_2)E_L(-\omega_1)} \ell_{eff}
		\end{equation}
	\end{subequations}
where N$_{bulk}$ is the bulk number density, and N$_{surf}$ is the surface number density (10$^{19}$ m$^{-2}$).\cite{RN133, RN503}	
For liquids with molar masses less than $\sim$100 g mol$^{-1}$ and densities roughly equal to that of H$_2$O$_{(l)}$ at room temperature, N$_{bulk} \sim$ 10$^{28}$ m$^{-3}$.
We take the bulk optical interaction length ($\ell_{eff}$) to be 10 $\mu$m.\cite{RN133} %see chapter 25 of Shen's NLO book... idk something here seems wrong
A normal dispersion curve is assumed so that $F(\omega_1+\omega_2) \approx F(-\omega_1+\omega_2 +\omega_3)$.
To simplify analysis, orientational averaging is ignored (e.g., $\langle \beta_{ijk} \mu_{l} \rangle \approx \beta \mu$) and the input and output fields are taken to be co-polarized, so that
\begin{equation}
		P_{ratio} \equiv \frac{\abs{P^{(3)}_{\text{HDFG}}}}{\abs{P^{(2)}_{\text{SFG}}}} \approx \frac{N_{bulk}}{N_{surf}} \frac{\beta}{\alpha} E(\omega_3) \ell_{eff} \sim 10^4 \frac{\beta}{\alpha} E(\omega_3)\\
\end{equation}
Ziegler has noted that for a field of intensity 10 GW/cm$^{2}$, $\frac{\beta E}{\alpha} \sim 10^{-3} $ for vibrational modes when $E(\omega_3)$ is largely detuned from electronic resonances. \cite{RN515}
Such an intensity is easily obtained using modern ultrafast sources.
In this limit, $P_\text{ratio} \sim 10$.
Since the intensity ratio scales as $\abs{P_{ratio}}^2$, we see that the HDFG output is somewhat stronger than vibrational SFG, assuming only interfacial contributions in vSFG.
It is important to note that this derivation ignored absorption and orientational averaging effects, which can significantly reduce output from either process.

@@

The steps behind orientational averaging of $\gamma_{ijkl}$, a rank four tensor in the molecular frame, are detailed elsewhere.\cite{Andrews1977, McDonnell2024}
Orientational averaging shows specific polarization schemes isolate linear combinations of different $\beta_{ijk}$ terms. 
Infrared spectra can provide $\abs{\vec{\mu}}$ values.
Since $\vec{\mu}$ is a rank-one tensor, its three Cartesian components share the simple relationship $\langle {\mu^2_x} \rangle = \langle {\mu^2_y} \rangle = \langle {\mu^2_z} \rangle = \frac{1}{3}\langle {\mu^2} \rangle$. \cite{RN459}
Neglecting rotational dependence (i.e., assuming only vibrational transitions), $\langle {\mu^2} \rangle = \mu^2$, so that $\langle \beta_{ijk} \mu_l \rangle = \langle \beta_{ijk} \rangle \frac{\abs{\mu}}{\sqrt{3}}$.
By introducing the reduced mass coordinate $q_n = \sqrt{\frac{1}{m_n}} Q_n$ and using the expansions of $\beta_{ijk}$ and $\mu_{l}$ to $\order{Q_n}$ found earlier, we see
\begin{equation}\label{betasive}
	\langle \frac{\partial \beta_{ijk}}{\partial q_n} \rangle = \varepsilon_0 \frac{48 \sqrt{3} }{NF}  \frac{\Gamma_{gv} \omega_{vg}}{\rho_{gg}} \frac{\Im(\chi^{(3)}_{IJKL})}{\frac{\partial \mu}{\partial q_n}}
\end{equation}
Since $\abs{\mu}$ values can be readily extracted from FT-IR spectra,\cite{RN412} this shows that HDFG can give quantitative information on the magnitude of $\beta_{ijk}$.
HDFG and vSFG together provide unique methods to extract the microscopic quantities ($\mu, \alpha, \beta$) that drive nonlinear output.
